\section*{2014}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{1}{a}{10} Describe a CUSUM control chart. Compare this chart with a Shewart chart with respect to performance. How is a V-mask useful in CUSUM charts?
\myline
\ques{2}{a}{15} Explain the concept of sequential samplin plan for attributes and obtain the average sampling umber function of a sequential sampling plan for testing the hypothesis \\
$H_0: \beta = \beta_0$ against the alternative hypothesis\\
$H_1: \beta = \beta_1$ in sampling from a population with density function $f(x,\beta)$.
\myline
\ques{3}{b}{20} Describe the principles on which a control chart is based. What is the concept behind the use of $3-\sigma$ liits is statistical quality control?\\\\
Control on measurements of pitch diameter threads in aircraft fittings is checked with 5 successive items measured at regular intervals. 8 such samples are given below:
\begin{center}
    \begin{tabular}{| *{6}{p{1.5cm}|}}
    \multicolumn{6}{l}{(Values are expressed in units of 0.0001 inch)} \\ \hline
    \emph{Sample} & \multicolumn{5}{p{8cm}|}{\emph{Measurement of each item of 5 items per hour}} \\ \hline
    1& 46 & 45 & 44 & 43 & 42 \\
    2& 41 & 41 & 45 & 42 & 41 \\
    3& 40 & 40 & 43 & 42 & 40 \\
    4& 42 & 43 & 43 & 42 & 45 \\
    5& 42 & 44 & 47 & 47 & 45 \\
    6& 39 & 46 & 45 & 42 & 41 \\
    7& 46 & 44 & 41 & 40 & 45 \\
    8& 40 & 45 & 43 & 40 & 39 \\ \hline
    \end{tabular}
\end{center}
Construct the $\overline{X}$ and $R$ chart and comment on process control.
\myline
\ques{4}{b}{10} Explain the terms
\begin{tasks}[after-item-skip = 0ex](3)
\task LTPD
\task Consumer's risk
\task Producers' risk
\task OC curve
\task Acceptance quality level
\end{tasks}
\ques{4}{c}{20} Suppose that the purchaser and the seller agree that a defective rate of 10\% in a lot is satisfactory. For the sampling plan $(N =20, n= 5, c=1)$, determine the probability that a lot with this percent (i.e. 10\%) defective will be accepted. Also determine the said probability if the sampling plan happens to be $N=100, n=5,c=1$. Which of these two sampling plans seems better and why?