\section*{2014}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{1}{b}{10} Let $ \{ X_n, \: n = 0,1,2, \ldots \}$ be a Markov chain on the state space $\{a , b , c\}$ with initial probability values $\displaystyle q_a = \frac{2}{5}, q_b = \frac{1}{5}, q_c = \frac{2}{5}$ and one step transition probability matrix $((p_{ij}))$ given by
\[ \left( \begin{array}{ccc}
    0 & \frac{1}{3} & \frac{2}{3}  \\
    \frac{1}{4} & \frac{3}{4} & 0 \\
    \frac{2}{3} & 0 & \frac{2}{3}
\end{array} \right) \]
Compute
\begin{enumerate}[itemsep = 0pt,label=(\roman*)]
    \item $P_r\left[X_1=b,X_2 =b,X_3 =b,X_4 =a,X_5 =c \middle|X_0 =a \right]$
    \item $P_r\left[X_1=b,X_2 =b,X_3 = a \right]$.
\end{enumerate}
\ques{1}{d}{10} To a single server queuing system, arrivals of customers are according to Poisson Process at the rate of 2 customers per hour. The service time has exponential distribution with mean service time of 12 minutes per customer. The steady state probability distribution $(p_n)$ for the number of customers $(n)$ is given below:
\begin{center}
    %\resizebox{10cm}{!}{
    \begin{tabular}{| c | *{6}{c|} c|}
    \hline
    $n$ & 0 & 1 & 2 & 3 & 4 & 5 & 6 and above \\ \hline
    $p_n$& 0.08 & 0.1 & 0.4 &0.2 & 0.2 & 0.02 & 0 \\ \hline
   \end{tabular}
\end{center}
Compute
\begin{enumerate}[ itemsep = -1ex,label=(\roman*)]
    \vspace{-1ex}
    \item the expected number of customers in the system and in the queue.
    \item the mean waiting time in the system and in the queue.
\end{enumerate}
\ques{3}{a}{15} For an $(M/M/1):(\infty/FIFO)$ queuing system, derive the set of steady-state difference equations and hence find the solution of the system. In a supermarket, the average arrival rate of customers is 5 every 30 minutes. The average time it takes to list and calculate the customers purchases at the cash desk is 4.5 minutes aand the time is exponentially distributed.
\begin{enumerate}[itemsep = -1ex,label=(\roman*)]
    \vspace{-1ex}
    \item How long will the customers expect to wait for service at the cash desk?
    \item What is the probability that the cashier is working?
\end{enumerate}